Stable individual differences in strategies within, but not between, visual search tasks

### 2019-02-22 Manuscript/manuscript.tex
Chopping more words!

\RequirePackage{fix-cm}
\documentclass[a4paper, jou, natbib]{apa6}

@@

\fiveaffiliations{Department of Psychology, University of Essex, Colchester, UK}{Department of Psychology, The Ohio State University, Columbus, USA}{School of Psychology, University of Aberdeen, Aberdeen, UK}{Department of Psychology, The Ohio State University, Columbus, USA}{School of Psychology, University of Aberdeen, Aberdeen, UK}
% {Corresponding author: a.clarke@essex.ac.uk}
% \renewcommand\Affilfont{\itshape\small}

% %%%% Subject entries to be placed here %%%%
% \subject{Behaviour, evolution}

% %%%% Keyword entries to be placed here %%%%
% \keywords{visual search, optimal behaviour, eye movements}

% %%%% Insert corresponding author and its email address}
% \corres{Alasdair Clarke\\
% \email{a.clarke@essex.ac.uk}}

@@

Here we focus on one source of individual differences in visual search: strategy. By strategy we mean a collection of search behaviours from which all observers can freely choose. Examples include adopting a systematic left-to-right and top-to-bottom strategy \citep{gilchrist2006}, or prioritizing locations that, based on knowledge or context, are more likely to contain the target \citep{wolfe2015}. A striking example of the effect of strategy is given by \cite{boot2006}. They asked participants to monitor a cluttered display for an object changing colour or suddenly appearing. Large individual differences were found with respect to the number of saccades participants made while monitoring the stimulus, which was negatively correlated with detection performance. In another example, Proulx \citeyearpar{proulx2011} demonstrated that participants differed in how they prioritised visual features (colour or orientation) when searching through a set of objects for one with a particular colour-orientation combination. These differences in search strategy explained some of the variation in the extent to which participants were distracted by an irrelevant object of a unique size. However, participants were largely unaware of the search strategy they employed.

Eye movement strategies have also been shown to be an important source of individual differences in visual search efficiency. \cite{nowakowska2017} designed a simple search paradigm to discriminate between optimal \citep{najemnik-geisler2008} and stochastic  \citep{clarke2016} search strategies. Participants searched through arrays of line segments (Figure \ref{fig:exampleStimuli}) arranged such that those on one side of the display all had a very similar orientation (homogeneous), while those on the other side had higher variance (heterogeneous). This meant that targets appearing on the homogeneous side were highly salient, while targets on the heterogeneous side were harder to find. The optimal eye movement strategy here is to search only the heterogeneous half, as targets on the homogeneous side can be detected with peripheral vision. We will refer to this paradigm here as the Split-Half Line Segment task (SHLS). Some participants searched the displays near optimally, but others carried out strategies counter to this, failing to even match the performance of the stochastic searcher. The degree to which participants made saccades in line with the optimal search strategy was strongly correlated with the speed of their search. A related version of this paradigm has been used in research investigating eye movement strategies in response to (simulated) hemianopia \citep{nowakowska2016,nowakowska2018}, with similar conclusions: the full spectrum of individual differences in strategy was observed, from near-optimal to the opposite of optimal. It is therefore not possible to conclude whether optimal or stochastic models better describe search without first explaining individual variability.

@@

Another example of differences in search strategy comes from the foraging literature  \citep{kristjansson2014,johannesson2016}. In this context, foraging tasks involve searching for multiple targets on each trial. Participants were asked to search through a set of items from four categories, with two categories classed as targets. In the conjunction condition (searching for red-horizontal and green-vertical line segments among red-vertical and green-horizontal distracters), most observers searched in runs, finding all the targets of one target category, and then switching and finding the targets in the other category. This strategy has previously been observed in animal foraging  \citep{dawkins1971}, and suggests holding one complex target template in mind at a time is a better strategy than switching templates. However, a sub-set of participants, termed `super-foragers', were able to change between search target categories with no cost to performance. This study has been run on touch displays \citep{kristjansson2014} (observers are asked to tap on as many targets as they can as quickly as possible) and a 3D computer game (in which participants navigate a squirrel through a virtual environment picking up targets with different feature combinations) \citep{prpic2018}, with super-foragers present in both versions. While test-retest reliability has not been measured explicitly for the foraging paradigm, the task was used as a measure to assess the effect of a six day long mindfulness retreat on cognitive performance \citep{hartkamp2017}. From a re-analysis of these data, we can estimate that the test-retest reliability for the mean run length is $r \approx 0.7$ for the feature condition and $r \approx 0.88$ for the conjunction search.

Previous research has investigated the relationship between these behaviours to psychometric measures, but to date, these differences have not shown strong correlations with other attributes. In the ACVS paradigm, Irons \& Leber found no evidence of a correlation between the proportion of optimal choices made by observers and measures of visual working memory; trait impulsivity \citep{irons-leber2016}; novelty seeking; need for cognition; and intolerance of uncertainty \citep{irons-leber2018}. Similarly, the differences in foraging behaviour are not accounted for by working memory or inhibitory control \citep{johannesson2017}.

A common theme emerging from these studies is the observation that individual strategies vary in their degree of effectiveness or optimality. However, ``visual search'' encompasses a wide range of specific tasks, each tapping into a different aspect of behaviour (e.g. feature-based attention, information sampling). The aim of the present study is to investigate the extent to which individual differences are stable across different visual search paradigms. Are observers who use the optimal strategy in the SHLS paradigm also more optimal in the ACVS task? Does it make sense to talk about `super-searchers' who show above-average performance in a range of search tasks (analogous to the 'super-recognizers' of the face-recognition literature  \citep{russell2009})? As a secondary question, we will measure the test-retest reliability of the differences found in the SHLS paradigm, and compare it with existing estimates of reliability for ACVS and MCFT.

@@

We view these findings not as a discouraging result, but as thought-provoking and exciting. Vogel and Awh \citep{vogel2008} argued that studying individual differences in cognitive psychology (in their case, working memory) provides valuable insight to containing potential theories of the underlying cognitive mechanisms. Our results suggest that context and structure of the task also needs to be taken into account. Understanding how an individual's behaviour varies across different search tasks can lead to the development of a more comprehensive theory of search.

>>>>

\RequirePackage{fix-cm}
\documentclass[a4paper, man, natbib, floatsintext]{apa6}

@@

\fiveaffiliations{Department of Psychology, University of Essex, Colchester, UK}{Department of Psychology, The Ohio State University, Columbus, USA}{School of Psychology, University of Aberdeen, Aberdeen, UK}{Department of Psychology, The Ohio State University, Columbus, USA}{School of Psychology, University of Aberdeen, Aberdeen, UK}

@@

Here we focus on one source of individual differences in visual search: strategy. By strategy we mean a collection of search behaviours from which all observers can freely choose. Examples include adopting a systematic left-to-right and top-to-bottom strategy \citep{gilchrist2006}, or prioritizing locations that, based on knowledge or context, are more likely to contain the target \citep{wolfe2015}. A striking example of the effect of strategy is given by \cite{boot2006}. They asked participants to monitor a cluttered display for an object changing colour or suddenly appearing. Large individual differences were found with respect to the number of saccades participants made while monitoring the stimulus, which was negatively correlated with detection performance.

Eye movement strategies have also been shown to be an important source of individual differences in visual search efficiency. \cite{nowakowska2017} designed a simple search paradigm to discriminate between optimal \citep{najemnik-geisler2008} and stochastic  \citep{clarke2016} search strategies. Participants searched through arrays of line segments (Figure \ref{fig:exampleStimuli}) arranged such that those on one side of the display all had a very similar orientation (homogeneous), while those on the other side had higher variance (heterogeneous). This meant that targets appearing on the homogeneous side were highly salient, while targets on the heterogeneous side were harder to find. The optimal strategy here is to search the heterogeneous half, as targets on the homogeneous side can be detected with peripheral vision. We will refer to this paradigm as the Split-Half Line Segment task (SHLS). Some participants searched the displays near optimally, but others carried out strategies counter to this, failing to even match the performance of the stochastic searcher. The degree to which participants made saccades in line with the optimal search strategy was strongly correlated with the speed of their search. A related version of this paradigm has been used in research investigating eye movement strategies in response to (simulated) hemianopia \citep{nowakowska2016,nowakowska2018}, with similar conclusions: the full spectrum of individual differences in strategy was observed. It is therefore not possible to conclude whether optimal or stochastic models better describe search without first explaining individual variability.

@@

Another example of differences in search strategy comes from the foraging literature  \citep{kristjansson2014,johannesson2016}. In this context, foraging tasks involve searching for multiple targets on each trial. Participants were asked to search through a set of items from four categories, with two categories classed as targets. In the conjunction condition (searching for red-horizontal and green-vertical line segments among red-vertical and green-horizontal distracters), most observers searched in runs, finding all the targets of one target category, and then switching and finding the targets in the other category. This strategy has previously been observed in animal foraging  \citep{dawkins1971}, and suggests holding one complex target template in mind at a time is a better strategy than switching templates. However, a sub-set of participants, termed `super-foragers', were able to change between search target categories with no cost to performance. While test-retest reliability has not been measured explicitly for the foraging paradigm, the task was used as a measure to assess the effect of a six day mindfulness retreat on cognitive performance \citep{hartkamp2017}. From a re-analysis of these data, we can estimate that the test-retest reliability for the mean run length is $r \approx 0.7$ for the feature condition and $r \approx 0.88$ for the conjunction search.

Previous research has investigated the relationship between these behaviours to psychometric measures, but to date, these differences have not shown strong correlations with other attributes. \cite{irons-leber2016,irons-leber2018} found no evidence of a correlation between the proportion of optimal choices made by observers in the ACVS paradigm and measures of visual working memory; trait impulsivity; novelty seeking; need for cognition; and intolerance of uncertainty. Similarly, the differences in foraging behaviour are not accounted for by working memory or inhibitory control \citep{johannesson2017}.

A common theme emerging from these studies is the observation that individual strategies vary in their degree of effectiveness or optimality. However, ``visual search'' encompasses a wide range of tasks, each tapping into a different aspect of behaviour (e.g. feature-based attention, information sampling). The aim of the present study is to investigate the extent to which individual differences are stable across different visual search paradigms. Does it make sense to talk about `super-searchers' who show above-average performance in a range of search tasks (analogous to the 'super-recognizers' of the face-recognition literature  \citep{russell2009})? As a secondary question, we will measure the test-retest reliability of the differences found in the SHLS paradigm, and compare it with existing estimates of reliability for ACVS and MCFT.

@@

We view these findings not as a discouraging result, but as thought-provoking and exciting. Vogel and Awh \citep{vogel2008} argued that studying individual differences in cognitive psychology (in their case, working memory) provides valuable insight to constraining potential theories of the underlying cognitive mechanisms. Our results suggest that context and structure of the task also needs to be taken into account. Understanding how an individual's behaviour varies across different search tasks can lead to the development of a more comprehensive theory of search.
